---
title: "2020-03-20 experiments"
output: 
  html_document: 
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
---

## Setup

### Load R libraries

```{r, message=FALSE}
library(tidyr)    # (Wickham & Henry, 2018)
library(ggplot2)  # (Wickham, 2009)
library(plyr)     # (Wickham, 2011)
library(dplyr)    # (Wickham et al., 2018)
library(cowplot)  # (Wilke, 2018)
```

## Load data

### Extract most performant organisms

```{r}

extract_max_fit <- function(data_path) {
  data <- read.csv(data_path, na.strings="NONE")
  data$matchbin_metric <- factor(data$matchbin_metric,
                                  levels=c("hamming", 
                                           "integer", 
                                            "integer-symmetric", 
                                            "hash", 
                                            "streak", 
                                            "streak-exact"))
  data$tag_mut_rate <- factor(as.numeric(data$MUT_RATE__FUNC_TAG_BF))
  return(data)
}

exp_name <- "2020-03-23"
dst_data_loc <- paste("../data/",exp_name,"/dir-sig/max_fit_orgs.csv", sep="")
dst_data <- extract_max_fit(dst_data_loc)
cst_data_loc <- paste("../data/",exp_name,"/chg-env/max_fit_orgs.csv", sep="")
cst_data <- extract_max_fit(cst_data_loc)

num_replicates <- 50
```

### Extract fitness over time

```{r}
dst_fot_data_loc <- paste("../data/", exp_name, "/dir-sig/fot.csv", sep="")
dst_fot_data <- read.csv(dst_fot_data_loc, na.strings="NONE")
dst_fot_data$tag_metric <- factor(dst_fot_data$tag_metric,
                                    levels=c("hamming", 
                                             "integer", 
                                             "integer-symmetric", 
                                             "hash", 
                                             "streak", 
                                             "streak-exact"))
dst_fot_data$tag_mut_rate <- factor(as.numeric(dst_fot_data$tag_mut_rate))

cst_fot_data_loc <- paste("../data/",exp_name,"/chg-env/fot.csv", sep="")
cst_fot_data <- read.csv(cst_fot_data_loc, na.strings="NONE")
cst_fot_data$matchbin_metric <- factor(cst_fot_data$matchbin_metric,
                                    levels=c("hamming", 
                                             "integer", 
                                             "integer-symmetric", 
                                             "hash", 
                                             "streak", 
                                             "streak-exact"))
cst_fot_data$tag_mut_rate <- factor(as.numeric(cst_fot_data$MUT_RATE__FUNC_TAG_BF))
```

## All Mutation Rates

### Solutions x Condition

Directional Signal Task

```{r}
generation_cutoffs <- c(500, 1000, 3000, 5000)
exp_prefix <- "dst"
for (gen in generation_cutoffs) {
  ggplot(filter(dst_data, solution=="1" & update <= gen), aes(x=tag_mut_rate, fill=tag_mut_rate)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ggtitle(paste("Directional signal task - solutions\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ylim(0, num_replicates+2) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-solutions-",gen,".png",sep=""), width=16, height=8)
}
```

Changing Signal Task

```{r}
generation_cutoffs <- c(50, 100, 500, 1000, 3000, 5000)
exp_prefix <- "cst"
for (gen in generation_cutoffs) {
  ggplot(filter(cst_data, solution=="1" & update <= gen), aes(x=tag_mut_rate, fill=tag_mut_rate)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ggtitle(paste("Changing signal task - solutions\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ylim(0, num_replicates+2) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-solutions-",gen,".png",sep=""), width=16, height=8)
}
```

### Score x Condition

Directional Signal Task

```{r}
generation_cutoffs <- c(5000)
exp_prefix <- "dst"
for (gen in generation_cutoffs) {
  ggplot(dst_data, aes(x=tag_mut_rate, y=aggregate_score, fill=tag_mut_rate)) +
    geom_boxplot() +
    ggtitle(paste("Directional signal task - score\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-scores-",gen,".png",sep=""), width=16, height=8)
}
```

Changing Signal Task

```{r}
generation_cutoffs <- c(5000)
exp_prefix <- "cst"
for (gen in generation_cutoffs) {
  ggplot(cst_data, aes(x=tag_mut_rate, y=score, fill=tag_mut_rate)) +
    geom_boxplot() +
    ggtitle(paste("Changing signal task - score\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-scores-",gen,".png",sep=""), width=16, height=8)
}
```

### Time to solution x Condition

```{r}
generation_cutoffs <- c(5000)
exp_prefix <- "dst"
for (gen in generation_cutoffs) {
  ggplot(dst_data, aes(x=tag_mut_rate, y=update, fill=tag_mut_rate)) +
    geom_boxplot() +
    ggtitle(paste("Directional signal task - time to solution\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-time--to-solution-",gen,".png",sep=""), width=16, height=8)
}
```

```{r}
generation_cutoffs <- c(5000)
exp_prefix <- "cst"
for (gen in generation_cutoffs) {
  ggplot(cst_data, aes(x=tag_mut_rate, y=update, fill=tag_mut_rate)) +
    geom_boxplot() +
    ggtitle(paste("Changing signal task - time to solution\n","<= ", gen, sep="")) +
    facet_wrap(~ matchbin_metric) +
    ggsave(paste("./imgs/",exp_name,"/", exp_prefix, "-time--to-solution-",gen,".png",sep=""), width=16, height=8)
}
```

## Best mutation rates

Mutation rates as determined by Matthew's experiments:

- Hamming: (2.00 / 1024) = 0.001953125
  - Expectation: [0.001:0.002]
- Hash: (0.75 / 1024) = 0.000732421875
  - Expectation: [0.0005:0.001]
- Integer: (6.00 / 1024) = 0.005859375
  - Expectation: [0.005:0.01]
- Integer (bi): (4.00 / 1024) = 0.00390625
  - Expectation: [0.002:0.005]
- Streak: (1.50 / 1024) = 0.00146484375
  - Expectation: [0.001:0.002]
  
### Directional signal task

```{r}
mut_rates <- levels(factor(dst_data$tag_mut_rate))
metrics <- levels(factor(dst_data$matchbin_metric))
for (metric in metrics) {
  print(paste("------------", metric, "------------"))
  for (mut_rate in mut_rates) {
    print(paste("Metric:", metric, "Mutation rate:", mut_rate))
    scores <- filter(dst_data,
                         matchbin_metric==metric & 
                           tag_mut_rate==mut_rate)$aggregate_score
    print(summary(scores))
    print("=======")
  }
}
```

Mutation rate results:

- Hamming
  - By solution count: 0.001
  - By median: 5e-04, 0.001, 0.002
  - By mean: 0.002
- Hash
  - By solution count: 0.001
  - By median: 0.001
  - By mean: 0.001
- Integer
  - By solution count: 0.002
  - By median: 5e-04, 0.002, 0.005
  - By mean: 0.005
- Integer-symmetric
  - By solution count: 0.0002
  - By median: 5e-04, 0.002, 0.005, 0.01
  - By mean: 0.01
- Streak
  - By solution count: 0.001, 0.002
  - By median: 2e-04, 5e-04, 0.001, 0.002
  - By mean: 0.002
  
### Changing-signal task

```{r}
mut_rates <- levels(factor(cst_data$tag_mut_rate))
metrics <- levels(factor(cst_data$matchbin_metric))
for (metric in metrics) {
  print(paste("------------", metric, "------------"))
  for (mut_rate in mut_rates) {
    print(paste("Metric:", metric, "Mutation rate:", mut_rate))
    scores <- filter(cst_data,
                         matchbin_metric==metric & 
                           tag_mut_rate==mut_rate)$score
    print(summary(scores))
    print("=======")
  }
}
```
