---
title: "2020-03-27 experiments"
output: 
  html_document: 
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
---

## Overview

Here, we compare the time required for solutions to evolve for the directional- and changing-signal tasks. We use identical configurations (with new random number seeds) as in the 2020-03-23-step-two experiments. We selected an _a priori_ threshold number of replciates to compare of 50% (or 100 out of 200 total replicates).

For these analyses, we only compared streak and hamming metrics.

## Setup

### Load R libraries

```{r, message=FALSE}
library(tidyr)    # (Wickham & Henry, 2018)
library(ggplot2)  # (Wickham, 2009)
library(plyr)     # (Wickham, 2011)
library(dplyr)    # (Wickham et al., 2018)
library(cowplot)  # (Wilke, 2018)

# Configure our default graphing theme
theme_set(theme_cowplot())
```

## Load data

### Extract most performant organisms

```{r}
extract_max_fit <- function(data_path) {
  data <- read.csv(data_path, na.strings="NONE")
  data$matchbin_metric <- factor(data$matchbin_metric,
                                  levels=c("hamming", 
                                           "integer", 
                                            "integer-symmetric", 
                                            "hash", 
                                            "streak", 
                                            "streak-exact"))
  data$tag_mut_rate <- factor(as.numeric(data$MUT_RATE__FUNC_TAG_BF))
  return(data)
}

exp_name <- "2020-03-27"
out_dir <- "2020-03-27"
dst_data_loc <- paste("../data/",exp_name,"/dir-sig/max_fit_orgs.csv", sep="")
dst_data <- extract_max_fit(dst_data_loc)
cst_data_loc <- paste("../data/",exp_name,"/chg-env/max_fit_orgs.csv", sep="")
cst_data <- extract_max_fit(cst_data_loc)

num_replicates <- 200
```

## Directional-signal Task Analyses

### How many replicates yielded a solution?

```{r}
generation_cutoffs <- c(500, 1000, 3000, 5000, 10000)
exp_prefix <- "dst"
for (gen in generation_cutoffs) {
  p <- ggplot(filter(dst_data, solution=="1" & update <= gen), aes(x=matchbin_metric, fill=matchbin_metric)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ggtitle(paste("DST solutions prior to generation ", gen, sep="")) +
    ylab("# successful replicates") +
    ylim(0, num_replicates + 2) +
    scale_x_discrete(name="Tag-matching metric",
                     limits=c("hamming",
                              "streak")) +
    theme(legend.position="none",
          axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/",out_dir,"/", exp_prefix, "-solutions-",gen,".png",sep=""), width=16, height=8)
  print(p)
}
```

### How many generations were required for solutions to evolve?

First we naively look at _all_ replicates (assuming replicates with no solution produced a solution in the final generation).

```{r}
exp_prefix <- "dst"
p <- ggplot(dst_data, aes(x=matchbin_metric, y=update, fill=matchbin_metric)) +
    geom_boxplot() +
    geom_jitter() +
    ggtitle(paste("Directional signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/",out_dir,"/", exp_prefix, "-time-to-solution.png",sep=""), width=16, height=8)
print(p)

ggplot(dst_data, aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    # geom_boxplot() +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, fun.args=list(conf.int=0.95), geom="errorbar", color="black") +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    ggtitle(paste("Directional signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8))
```

Next, we do a fair comparision between first 100 successful replicates of each metric.

```{r}
metrics <- c("hamming", "hash", "integer", "integer-symmetric", "streak")
dst_update_rankings <- data.frame(matchbin_metric=character(), 
                                  aggregate_score=numeric(), 
                                  update=numeric(),
                                  update_rank=numeric(),
                                  solution=numeric())
for (metric in metrics) {
  dst_metric <- 
    filter(dst_data, matchbin_metric==metric)
  dst_metric <- 
    subset.data.frame(dst_metric, select=c("aggregate_score", 
                                           "update", 
                                           "matchbin_metric",
                                           "solution"))
  dst_metric$update <- as.numeric(dst_metric$update)
  dst_metric$update_ranking <- rank(dst_metric$update, ties.method="random")
  dst_update_rankings <- rbind(dst_update_rankings, dst_metric)
}

threshold <- 100
exp_prefix <- "dst"
ggplot(filter(dst_update_rankings, update_ranking <= threshold), aes(x=matchbin_metric, y=update, fill=matchbin_metric)) +
    geom_jitter(aes(color=matchbin_metric)) +
    geom_boxplot(alpha=0.75) +
    ggtitle(paste("Directional signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/", out_dir,"/", exp_prefix, "-time-to-solution-top-",threshold,"-bp.png", sep=""))

ggplot(filter(dst_update_rankings, update_ranking <= threshold), aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, fun.args=list(conf.int=0.95), geom="errorbar", color="black") +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    ggtitle(paste("Directional signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/", out_dir,"/", exp_prefix, "-time-to-solution-top-",threshold,"-ci.png", sep=""))

```

```{r}
dst_first_k_solutions <- filter(dst_update_rankings, update_ranking <= threshold)
wilcox.test(formula = update ~ matchbin_metric, data=dst_first_k_solutions,
                     exact=FALSE,
                     conf.int=TRUE,
                     conf.level=0.95)
```

Indeed, there is a significant difference in time to solution.

## Changing-signal Task Analyses

### How many replicates yielded a solution?

```{r}
generation_cutoffs <- c(50, 100, 500)
exp_prefix <- "cst"
for (gen in generation_cutoffs) {
  p <- ggplot(filter(cst_data, solution=="1" & update <= gen), aes(x=matchbin_metric, fill=matchbin_metric)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ggtitle(paste("Changing signal task - solutions\n","<= ", gen, sep="")) +
    ylim(0, num_replicates+2) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/",out_dir,"/", exp_prefix, "-solutions-",gen,".png",sep=""), width=16, height=8)
  print(p)
}
```
### How many generations were required for solutions to evolve?

First, we naively look at all replicates, assuming that unsuccessful replicates produced solutions in the final generation.

```{r}
exp_prefix <- "cst"
p <- ggplot(cst_data, aes(x=matchbin_metric, y=update, fill=matchbin_metric)) +
    geom_boxplot() +
    geom_jitter() +
    ggtitle(paste("Changing signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/",out_dir,"/", exp_prefix, "-time-to-time-to-solution.png",sep=""), width=16, height=8)
print(p)

ggplot(cst_data, aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, fun.args=list(conf.int=0.95), geom="errorbar", color="black") +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    ggtitle(paste("Changing signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8))
```

Next we do a fair comparison looking at first 50% of successful replicates.

```{r}
metrics <- c("hamming", "hash", "integer", "integer-symmetric", "streak")
cst_update_rankings <- data.frame(matchbin_metric=character(), 
                                  score=numeric(), 
                                  update=numeric(),
                                  update_rank=numeric(),
                                  solution=numeric())
for (metric in metrics) {
  cst_metric <- 
    filter(cst_data, matchbin_metric==metric)
  cst_metric <- 
    subset.data.frame(cst_metric, select=c("score", 
                                           "update", 
                                           "matchbin_metric",
                                           "solution"))
  cst_metric$update <- as.numeric(cst_metric$update)
  cst_metric$update_ranking <- rank(cst_metric$update, ties.method="random")
  cst_update_rankings <- rbind(cst_update_rankings, cst_metric)
}

threshold <- 100
exp_prefix <- "cst"
ggplot(filter(cst_update_rankings, update_ranking <= threshold), aes(x=matchbin_metric, y=update, fill=matchbin_metric)) +
    geom_jitter(aes(color=matchbin_metric)) +
    geom_boxplot(alpha=0.75) +
    ggtitle(paste("Changing signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/", out_dir,"/", exp_prefix, "-time-to-solution-top-",threshold,"-bp.png", sep=""))

ggplot(filter(cst_update_rankings, update_ranking <= threshold), aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, fun.args=list(conf.int=0.95), geom="errorbar", color="black") +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    ggtitle(paste("Changing signal task - time to solution", sep="")) +
    theme(axis.text.x=element_text(size=8)) +
    ggsave(paste("./imgs/", out_dir,"/", exp_prefix, "-time-to-solution-top-",threshold,"-ci.png", sep=""))

```

```{r}
cst_first_k_solutions <- filter(cst_update_rankings, update_ranking <= threshold)
wilcox.test(formula = update ~ matchbin_metric, data=cst_first_k_solutions,
                     exact=FALSE,
                     conf.int=TRUE,
                     conf.level=0.95)
```

No significant difference between hamming and streak metrics.

## Paper figure

Here's the source code for generating the figure in the paper.

```{r}

all_sol_data_dst <- "../data/2020-03-23/step-two/dir-sig/max_fit_orgs.csv"
all_sol_data_cst <- "../data/2020-03-23/step-two/chg-env/max_fit_orgs.csv"
dst_data_03_23 <- extract_max_fit(all_sol_data_dst)
cst_data_03_23 <- extract_max_fit(all_sol_data_cst)

dst_sols_p <- 
  ggplot(filter(dst_data_03_23, solution=="1" & update <= 5000), 
         aes(x=matchbin_metric, fill=matchbin_metric)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ylab("Solutions") +
    ylim(0, num_replicates + 4) +
    scale_x_discrete(name="Metric",
                     limits=c("hamming",
                              "integer",
                              "integer-symmetric",
                              "hash",
                              "streak"),
                     labels=c("Hamming",
                              "Integer",
                              "Integer (bi)",
                              "Hash",
                              "Streak")) +
    theme(legend.position="none",
          axis.text.x=element_text(size=8)) 

cst_sols_p <- 
  ggplot(filter(cst_data_03_23, solution=="1" & update <= 100), 
         aes(x=matchbin_metric, fill=matchbin_metric)) +
    geom_bar(stat="count", position="dodge") +
    geom_text(stat="count", 
            mapping=aes(label=..count..), 
            position=position_dodge(0.9), vjust=0) +
    ylab("Solutions") +
    ylim(0, num_replicates + 10) +
    scale_x_discrete(name="Metric",
                     limits=c("hamming",
                              "integer",
                              "integer-symmetric",
                              "hash",
                              "streak"),
                     labels=c("Hamming",
                              "Integer",
                              "Integer (bi)",
                              "Hash",
                              "Streak")) +
    theme(legend.position="none",
          axis.text.x=element_text(size=8)) 

cst_time_p <- 
  ggplot(filter(cst_update_rankings, update_ranking <= threshold), 
         aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, 
                 fun.args=list(conf.int=0.95), 
                 geom="errorbar", 
                 color="black") +
    scale_x_discrete(name="Metric",
                     limits=c("hamming",
                              "streak"),
                     labels=c("Hamming",
                              "Streak")) +
    ylim(0, 100) +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    theme(legend.position = "none",
          axis.text.x=element_text(size=8)) 

dst_time_p <- 
  ggplot(filter(dst_update_rankings, update_ranking <= threshold), 
         aes(x=matchbin_metric, y=update, color=matchbin_metric)) +
    geom_jitter(alpha=0.75) +
    stat_summary(fun.data=mean_cl_boot, 
                 fun.args=list(conf.int=0.95), 
                 geom="errorbar", 
                 color="black") +
    scale_x_discrete(name="Metric",
                     limits=c("hamming",
                              "streak"),
                     labels=c("Hamming",
                              "Streak")) +    ylim(0, 5000) +
    stat_summary(fun.y = mean, geom = "point", colour = "black") +
    theme(legend.position = "none",
          axis.text.x=element_text(size=8)) 


( ( cst_sols_p | cst_time_p ) / ( dst_sols_p | dst_time_p ) ) + plot_annotation(tag_levels='A') + ggsave("panel.pdf")

```
